---
title: "HW10"
output: html_document
date: "2025-05-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### Downloading the dataset from TidyTuesday. This Dataset incluedes information about an animal shelter in Florida. 
```{r, eval=TRUE,echo=TRUE}
#libraries
library(ggplot2)
library(ggthemes)
library(patchwork)
library(waffle)
library(ggridges)
library(ggbeeswarm)
library(GGally)
library(ggpie)
library(ggmosaic)
library(scatterpie)
library(DescTools)
library(treemap)
library(tidyverse)
library(hrbrthemes)
library(beeswarm)
library(treemapify)

#Dataset
longbeach <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-04/longbeach.csv')

#cleaning script from the TigyTuesday
library(dplyr)
#install.packages("animalshelter")
library(animalshelter)

longbeach <- animalshelter::longbeach |>
  dplyr::mutate(
    was_outcome_alive = as.logical(was_outcome_alive),
    dplyr::across(
      c(
        "animal_type",
        "primary_color",
        "secondary_color",
        "sex",
        "intake_condition",
        "intake_type",
        "intake_subtype",
        "reason_for_intake",
        "jurisdiction",
        "outcome_type",
        "outcome_subtype"
      ),
      as.factor
    )
  ) |>
  dplyr::select(-"intake_is_dead")

#checking the variables
str(longbeach)
```
#### 1.Waffle plot to see the proportions of animal type in the shalter 
```{r, eval=TRUE,echo=TRUE}
# Prepare data (Example: Count property types)
waffle_data <- longbeach %>%
  count(animal_type) %>%
  mutate(proportion = n / sum(n) * 100)

# Creating the waffle plot
ggplot(waffle_data, aes(fill=animal_type, values=proportion)) +
  geom_waffle() +
  theme_void()
```
#### 2. Pie chart about the sex of the animals
```{r, eval=TRUE,echo=TRUE}
pie_data<-longbeach %>%
  count(sex)%>%
  mutate(proportion = `n` / sum(`n`))%>%
  mutate(labels = scales::percent(proportion))

#creating the pie chart
ggplot(pie_data, aes(x = "", y = proportion, fill = sex)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels),
            position = position_stack(vjust = 0.5),
            show.legend = FALSE) +
  guides(fill = guide_legend(title = "Sex")) +
  scale_fill_brewer(palette="Set1") +
  coord_polar(theta = "y") + 
  theme_void()
```
#### 3. Treemap with information about the reason for animal intake such as stray capture, wildlife captures, adopted but returned, owner surrendered etc.
```{r, eval=TRUE,echo=TRUE}
tree_data<-longbeach %>%
  count(intake_type)%>%
  mutate(proportion = `n` / sum(`n`))%>%
  mutate(labels = scales::percent(proportion))

ggplot(tree_data, aes(area = n, fill =intake_type)) +
  geom_treemap()
```
#### 4.
```{r, eval=TRUE,echo=TRUE}
ggplot(tree_data, aes(x=intake_type, y=n, fill=intake_type)) + # fill=name allow to automatically dedicate a color for each group
  geom_violin()
```
