---
title: "Homework 5"
output: html_document
date: "2025-02-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 1. Assign to the variable n_dims a single random integer between 3 and 10
```{r, eval=TRUE,echo=TRUE}
n_dims <- runif(1, min = 3, max = 11)
print(n_dims)
```
##### Create a vector of consecutive integers from 1 to n_dims^2^
```{r, eval=TRUE,echo=TRUE}
vec_con<-1:(n_dims^2)
print(vec_con)
```

##### Use the sample function to randomly reshuffle these values.
```{r, eval=TRUE,echo=TRUE}
vec_samp<-sample(vec_con)
print(vec_samp)
```
##### Create a square matrix with these elements.
```{r, eval=TRUE,echo=TRUE}
m <- matrix(vec_samp, nrow = n_dims, ncol = n_dims)
print(m)
```
##### Find a function in r to transpose the matrix.
```{r, eval=TRUE,echo=TRUE}
transposed_matrix<-t(m)
print(transposed_matrix)
```
##### Calculate the sum and the mean of the elements in the first row and then the last row.
```{r, eval=TRUE,echo=TRUE}
sum_first_row <- sum(transposed_matrix[1, ])
print(sum_first_row)
mean_first_row <- mean(transposed_matrix[1, ])
print(mean_first_row)
sum_last_row <- sum(transposed_matrix[n_dims, ])
print(sum_last_row)
mean_last_row <- mean(transposed_matrix[n_dims, ])
print(mean_last_row)
```
##### Read about the eigen() function and use it on your matrix
```{r, eval=TRUE,echo=TRUE}
eigen_matrix <- eigen(transposed_matrix)
print(eigen_matrix$values)
print(eigen_matrix$vectors)
```
##### Dig in with the typeof() function to figure out their type.
```{r, eval=TRUE,echo=TRUE}
typeof(eigen_matrix$values)
typeof(eigen_matrix$vectors)
```


#### 2. Create a list with the following named elements:


##### A 4 x 4 matrix filled with random uniform values
```{r, eval=TRUE,echo=TRUE}
my_matrix <- matrix(runif(16), nrow = 4, ncol = 4)
print(my_matrix)
```
##### A 100-element vector of TRUE or FALSE values
```{r, eval=TRUE,echo=TRUE}
random_values <- runif(100)
my_logical <- random_values > 0.5
print(my_logical)
```
##### A 26-element vector of all the lower-case letters in random order.
```{r, eval=TRUE,echo=TRUE}
my_letters <- sample(letters)
print(my_letters)
```
##### The list
```{r, eval=TRUE,echo=TRUE}
my_list <- list(my_matrix = my_matrix, my_logical = my_logical, my_letters = my_letters)
```
##### Create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
```{r, eval=TRUE,echo=TRUE}
my_list2 <- list(my_matrix[2, 2], my_logical[2], my_letters[2])
print(my_list2)
```
##### Use the typeof() function to confirm the underlying data types of each component in this list
```{r, eval=TRUE,echo=TRUE}
sapply(my_list2, typeof)
```
##### Combine the underlying elements from the new list into a single atomic vector with the c() function.
```{r, eval=TRUE,echo=TRUE}
combined_vec <- c(my_list[[1]], my_list[[2]], my_list[[3]])
```
##### What is the data type of this vector?
```{r, eval=TRUE,echo=TRUE}
typeof(combined_vec)
```

#### 3.Create a data frame with the two variables (= columns) and 26 cases (= rows) below:


##### Call the first variable my_unis and fill it with 26 random uniform values from 0 to 10
```{r, eval=TRUE,echo=TRUE}
my_unis <- runif(26, min = 0, max = 10)
```
##### Call the second variable my_letters and fill it with 26 capital letters in random order.
```{r, eval=TRUE,echo=TRUE}
my_letters <- sample(letters)
```
##### Create the data frame
```{r, eval=TRUE,echo=TRUE}
my_data_frame <- data.frame(my_unis = my_unis, Letter = my_letters)
print(my_data_frame)
```
##### For the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA
```{r, eval=TRUE,echo=TRUE}
my_data_frame$my_unis[sample(1:26, 4)] <- NA
```
##### For the first variable, write a single line of R code to identify which rows have the missing values.
```{r, eval=TRUE,echo=TRUE}
which(is.na(my_data_frame$my_unis))
```
##### Re-order the entire data frame to arrange the second variable in alphabetical order
```{r, eval=TRUE,echo=TRUE}
my_data_frame <- my_data_frame[order(my_data_frame$Letter), ]
print(my_data_frame)
```
##### Calculate the column mean for the first variable.
```{r, eval=TRUE,echo=TRUE}
mean_my_unis <- mean(my_data_frame$my_unis, na.rm = TRUE)
print(mean_my_unis)
```

