---
title: "HW6"
output: html_document
date: "2025-02-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### Creating Fake Data Sets To Explore Hypotheses
##### I'm creating a fake data set on the effects of dolphin watching activities on dolphin behaviour. This can be done by quantifying the occurrence of avoidance responses, one of which is a change in the dolphin's swimming direction. In this homework I will test the occurrence of directional changes in two conditions: no tourist boats present (control) and with tourist boats present. Based on this, I hypothesise that dolphins will show a greater number of directional changes in the presence of tourist boats.
```{r, eval=TRUE,echo=TRUE}
# Specify parameters
n_no_boats <- 50  # Sample size for the group without boats (number of dolphins)
mean_no_boats <- 9  # Mean of occurrences of directional changes in the group without boats
var_no_boats <- 10   # variance for the group without boats

n_boats <- 50  # Sample size for the group with boats
mean_boats <- 20  # Mean of occurrences of directional changes in the group with boats
var_boats <- 15  # variance for the group with boats

# Generate random data
No_boats <- round(rnorm(n_no_boats, mean = mean_no_boats, sd = sqrt(var_no_boats)),0)
boats <- round(rnorm(n_boats, mean = mean_boats, sd = sqrt(var_boats)),0)

# Combine data into a data frame
data <- data.frame(
  Directional_changes = c(No_boats, boats),
  Group = rep(c("No_boats", "boats"), each = n_no_boats)
)
head(data)
str(data)
```
##### Analyzing the data
```{r, eval=TRUE,echo=TRUE}
library(ggplot2)
#ANOVA analysis
anova_result <- aov(Directional_changes ~ Group, data = data)
summary(anova_result)

# Generate a boxplot to visualize the data
library(ggplot2)

ggplot(data, aes(x = Group, y = Directional_changes, fill = Group)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    x = "Group",
    y = "Number of directional changes"
  ) 
```
#### Using a for loops to adjust the parameters and explore how they might impact the results/analysis.How small can your sample size be before you detect a significant pattern (p < 0.05)? How small can the differences between the groups be (the “effect size”) for you to still detect a significant pattern?
```{r, eval=TRUE,echo=TRUE}
library(dplyr)
#variables to explore
effect_sizes <- seq(1, 10, 0.5)  # Varying effect sizes
sample_sizes <- seq(5, 50, 5) # Varying sample sizes
alpha <- 0.05

# data frame to store the results
sample_size_results <- data.frame()
# creating the loop
for (effect in effect_sizes) {
  for (n in sample_sizes) {
    group_boats <- rnorm(n, mean = mean_boats, sd = sqrt(var_boats))
    group_no_boats <- rnorm(n, mean = mean_no_boats + effect, sd = sqrt(var_no_boats))
    t_test <- t.test(group_boats, group_no_boats)
    p_value <- t_test$p.value
    # storing the results
    sample_size_results <- rbind(sample_size_results, 
                     data.frame(Effect_Size = effect, 
                                Sample_Size = n, 
                                P_Value = p_value, 
                                Significant = (p_value < alpha)))
  }
}

# View the first few rows of results
print(head(sample_size_results))
# filtering the data frame for significant results with small sample sizes
sample_size_results %>% filter(Significant == TRUE & Sample_Size <= 5)
```
##### The table above shows that the minimum sample size to see significant differences between groups is 5, and a differences of 1 between means is enough to still see a difference. 
