---
title: "HW8"
output: html_document
date: "2025-03-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Simulating and Fitting Data Distributions
##### Running the code on my data. The variable chosen is the inter-breath intervals (dive times, in seconds) of bottlenose dolphins. These measurements were taken in the Mexican Caribbean between 2021 and 2022 (Paredes-Torres et al., 2025).
```{r, eval=TRUE,echo=TRUE}
library(ggplot2)
library(MASS)
# my dataset
z <- read.table("ibi.csv",header=TRUE,sep=",")
str(z)
summary(z)

## Plot histogram of data
p1 <- ggplot(data=z, aes(x=IBI, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",linewidth=0.2) 
print(p1)
##Get maximum likelihood parameters for normal
normPars <- fitdistr(z$IBI,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
##Plot normal probability density
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$IBI),len=length(z$IBI))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$IBI), args = list(mean = meanML, sd = sdML))
p1 + stat
##Plot exponential probability density
expoPars <- fitdistr(z$IBI,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$IBI), args = list(rate=rateML))
p1 + stat + stat2

##Plot uniform probability density
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$IBI), args = list(min=min(z$IBI), max=max(z$IBI)))
p1 + stat + stat2 + stat3

###Plot gamma probability density
gammaPars <- fitdistr(z$IBI,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$IBI), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4
##Plot beta probability density
pSpecial <- ggplot(data=z, aes(x=IBI/(max(IBI + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$IBI/max(z$IBI + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$IBI), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

#### Simulate a new data set. Using the best-fitting distribution, go back to the code and get the maximum likelihood parameters. Use those to simulate a new data set, with the same length as your original vector, and plot that in a histogram and add the probability density curve. Right below that, generate a fresh histogram plot of the original data, and also include the probability density curve.

##### The best-fitting distribution for my dataset seems to be the gamma distribution

```{r, eval=TRUE,echo=TRUE}
# Simulating random numbers using the rgamma function
sim <- rgamma(n=826,shape =shapeML,rate = rateML)
simu_data <- data.frame(1:826,sim)
names(simu_data) <- list("ID","IBI")
str(simu_data)
summary(simu_data$IBI)
#plot an histogram of the data 
p <- ggplot(data=simu_data, aes(x=IBI, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p)
# calculating the maximum likelihood parameters
gammaP <- fitdistr(simu_data$IBI,"gamma")
shapeMLsim <- gammaP$estimate["shape"]
rateMLsim <- gammaP$estimate["rate"]

stat_simgamma <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(simu_data$IBI), args = list(shape=shapeMLsim, rate=rateMLsim))
#adding the probability density curve
p+stat_simgamma
```

##### Now, ploting again the original data and with the gamma density curve
```{r, eval=TRUE,echo=TRUE}
p1+stat4
```

#### How do the two histogram profiles compare? Do you think the model is doing a good job of simulating realistic data that match your original measurements? Why or why not? 
##### I think the simulation is doing a good job, so the density curve adjuts better than in the original data.
